\section{Modeling Energy Systems}

Energy systems are spatially, temporally, and topologically complex ‘machines’
that operate cohesively to supply energy where it is needed at the exact moment
it is needed. Beyond the physical and engineering aspects of energy systems is a 
social context that motivates choices at every layer of complexity 
\cite{dotson_influence_2022}. \Acfp{esom} therefore have several possible purposes such as forecasting future quantities, generating insight for policy development, or energy system planning for scheduling and acquisition \cite{decarolis_using_2011, yue_review_2018}.
\textcolor{red}{SIREN \cite{sustainable_energy_now_renewable_2016} can actually do multi-objective optimization with a GUI. It's not as flexible as mine will be, but it \textit{can} do it.}

\textcolor{red}{EnergyScope is \textit{suitable} for uncertainty analysis (i.e. many runs are computationally tractable) but does not have any built-in capabilities \cite{limpens_energyscope_2019}.}

\begin{table}
    \centering
    \caption{Summary of \ac{esom} frameworks.}
    \label{tab:esoms}
    \resizebox*{\textwidth}{0.95\textheight}{\input{tables/esom_database.tex}}
\end{table}



These papers assume inelastic electricity demand
\begin{enumerate}
    \item Hueberger et al. 2017 \cite{heuberger_power_2017}
\end{enumerate}

These papers allow elastic electricity demand
\begin{enumerate}
    \item Leuthold et al. 2008 \cite{leuthold_elmod_2008}
\end{enumerate}

\begin{itemize}
    \item METIS has the following motivation \cite{sakellaris_metis_2018}
    \begin{enumerate}
        \item Close the gap between modellers and policy-makers, enabling policy-makers to become modellers.
        \item reconciles user-friendliness with powerful capabilities
        \item modularity
        \item \textcolor{red}{METIS does not support multi-objective optimization!}    
    \end{enumerate}
    \item energyRt has similar motivations \cite{lugovoy_energyrt_2022}.
    \begin{enumerate}
        \item enhance reproducibility
        \item reduce the learning curve
        \item minimize model development time
    \end{enumerate}
    \item GENESYS cannot \sout{can} do multi-objective optimization \cite{bussar_optimal_2014}. It uses a logical flow to model dispatch behavior for a genetic algorithm!
    \begin{enumerate}
        \item One drawback of the "hierarchical system management" method (i.e. a logical flow chart) is the difficulty to model ramping rates... (test this).
    \end{enumerate}
\end{itemize}

Some models such as MEDEAS \cite{capellan-perez_medeas_2020} and MultiMod \cite{huppmann_market_2014} are general equilibrium models which embed energy
systems within the macro-economy and facilitate modeling of strategic behavior 